---
title: "In-Class Computing Task"
author: "Math 253: Statistical Computing & Machine Learning"
date: "Tuesday Sept 22, 2015"
output: 
  rmarkdown::tufte_handout
---

```{r include=FALSE}
options(width=80)
library(mosaicData)
```

# Histogram function

## Task 1

Write a function `myHistogram()` that packages up the commands you wrote for drawing a histogram into a function.  It should draw the histogram (setting up a frame, as appropriate) and return the bin counts.

Give it an argument `fill` to set the color to fill in the bars.  Give this a nice default value.

In addition, your program should draw a normal (a.k.a. gaussian) density in red over the histogram.  The function to calculate density values is `dnorm()`.  You'll need to give it the mean and standard deviation of the data being binned, which you can calculate directly from those data.

# Density estimation

## Task 2

Work throught the following steps to generate a density plot. 


You've got a set of numbers `v`. You want to plot the density of `v`.
```{r}
v <- rnorm(10, mean=100, sd=1)
```



1. Pick a kernel wide enough to span the gaps between most points.  Call the width the *bandwidth* (`bw`).  A good starting value is the range divided by the square root of the number of values. Then create a function called `kernel()` that is a smooth density centered on any specified value `x`

```{r}
bw <- diff(range(v)) / sqrt(length(v))
kernel <- function(v, x) {dnorm(v, mean=x, sd=bw)}
```

2. Create a set of values `x` at which to evaluate the density.  This should extend several bandwidths to either side of the range of `v`. 
```{r}
x <- seq(min(v) - 5*bw, max(v) + 5*bw, length=200)
```

3. Now, for each of the values in `v`, for each of the values in `x`, calculate the value of your `kernel()` function.
```{r}
Dvals <- outer(v, x, FUN = kernel)
```
The result will be a matrix with `length(v)` rows and `length(x)` columns.

4. To convert this to a density, `density`, calculate the column sums of the matrix and divide by `length(v)`. (See `colSums()`.) You can check your result by confirming that the sum of `dens` divided by the interval between successive values of `x` gives 1. 

5. Produce a data frame with components `x` and `density`.

6. You can plot that data frame to see the density estimate.

# Task 3

Package your commands into a function, `plotdensity()` that takes `v` as an argument and produces a nice density plot.  Allow the user to override `xlim` by including it as an argument with a default value of `NULL`.

The function should return the data frame from step (5) above with `x` and `density`.  Since it's a graphics function, some people will want to avoid having to look at the return value; they will be interested just in the graphical *side effect*.  You can use `invisible()` in place of `return()` to accomplish this.

# Above and beyond

One of the nice features of R is the ability to use variables as arguments to functions without quoting them.  For instance:

```{r eval=FALSE}
plot(y ~ x, data=MyData)
```

You can arrange your `plotdensity()` function to work this way.  Here's a template without the details of drawing the density plot.

```{r}
plotdensity <- function(v, data=parent.frame(), ...) {
  v <- substitute(v) # a version of v that won't be evaluated until asked
  vname <- as.character(v)
  vvals <- eval(v, envir = data)
  # Your statements go here. 
  # Use vvals as the data and vname as the label for the x-axis
}
```
