---
title: "Style Faux Pas"
author: "Statistical Computing & Maching Learning"
date: "August 31, 2015"
output: html_document
---

# It's not T

The book sometimes uses a variable named `T` in a command like that on p.49:
```{r eval=FALSE}
Auto=read.csv("Auto.csv",header=T,na.strings="?")
```

You should not use `T` to mean `TRUE`.  `TRUE` is a keyword in R whose meaning can never change.  `T` is a garden-variety variable that might mean `TRUE` or might not.  Truth should be certain.  `TRUE` is, `T` isn't.

Similarly, use `FALSE` and not `F`.

Other ways in which the above command is a style faux pas.

1. Use `<-` not `=` for assignment.    
    There are two completely distinct operations in R.    
    1. Assignment of a value to an object name.  Use `<-` for this, even though it's allowed by R to use `=`.
    2. Identification of the name of an argument.  You must use `=` for this.
2. Use spacing to help the reader (i.e. yourself).      
    A couple of simple rules will get you far:
    * Put a single space after a `,` but not before.
    * Put a space on either side of `=` and `<-`.
    
Re-written to avoid faux pas, the statement becomes:
```{r eval=FALSE}
Auto <- read.csv("Auto.csv", header = T, na.strings = "?")
```

To give the authors of ISL due credit, there are important things about the above statement that is completely correct: it uses a leading capital letter (`Auto`) to name a data table and leading lower-case letters (`horsepower`) as variable names.

Using spaces in a consistent way helps you to distinguish between different parts of a command and let's you read the command in the same way that it will be executed.

# The horror of attachment

*An Introduction to Statistical Learning* makes use of a function called `assign()`.  This is a bad practice.

But it's not enough to say that.  It's important to know what are the good practices to be used instead.

## Example 1

On p. 50 of ISL, you'll see:

> To refer to a variable, we must type the data set and the variable name joined with a `$` symbol.  Alternatively, we can use the `attach()` function in order to tell R to make the variables in this data frame available by name.
```{r fig.keep='none', eval=FALSE}
plot(Auto$cylinders, Auto$mpg)
attach(Auto)
plot(cylinders, mpg)
```
```{r echo=FALSE, eval=FALSE}
detach(Auto)
```



## Being pedantic

Of all the kinds of objects there are in R, `assign()` is a `function`.
```{r}
class(assign)
```
But it is not a function in the mathematical sense.  A mathematical function takes one or more values as input and returns a value as output.  `Assign()` doesn't return anything.  Instead, it manipulates your R session in ways that are mysterious and unknown to the very people for whom `assign()` is most useful: beginners.

The documentation for `attach()` has a section labelled "Good practice." Read it and see if you think a beginner would be able to make any sense of it.  "Good practice" is not good practice unless users can use it!